---
title: "Statistical Inference Course Project"
author: "Javier Nieto"
date: "Monday, January 12, 2015"
output: pdf_document
---

# Part 2: Basic inferential data analysis

Simple inferential data analysis
---

Now in the second part of the project, we analyze the ToothGrowth data in the R 
datasets package and build a simple inferential data analysis.


Setting up required environment in R
---

```{r,results='hide'}
# Load libraries
library(ggplot2)
library(knitr)

# Changing locale time to English
Sys.setlocale("LC_TIME", "english")

```

Data
---

The data we use are ToothGrowth, which are the in the R datasets package. These 
data contain information about the effect of vitamin C on tooth growth in guinea
pigs.

The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at 
each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery
methods (orange juice or ascorbic acid).

Format

A data frame with 60 observations on 3 variables.

[,1]  len	numeric	Tooth length

[,2]  supp factor Supplement type (VC or OJ).

[,3]  dose numeric Dose in milligrams.


Source

C. I. Bliss (1952) The Statistics of Bioassay. Academic Press.


Exploratory data analysis
---

```{r}
require(graphics)
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth,
       xlab = "ToothGrowth data: length vs dose, given type of supplement")
```



Basic summary of the data
---


Inferential data analysis
---


Conclusions
---



